\section*{Conclusions and future work}
\quot{Poetry is a sword of lightning, ever unsheathed, \\which
  consumes the scabbard that would contain it.}{Percy Bysshe Shelley}

\paragraph{}{This paper has discussed translation of poetry in a
  computational, statistically-based way. A discussion of the history
  of machine translation, the theory and practice of modern machine
  translation systems, and current and possible approachs to manual
  and computational translation, have all allowed me to explore the
  possibility of reaching high quality, fully automatic machine
  translation of poetry. }
\paragraph{}{The hypotheses I have discussed, of how current machine
  translation methods may be improved to create better translations of
  poetry, all lead nicely into future work. Each of the methods I have
  proposed --- training the language model with a custom-built corpus,
  altering the alignment model, and recovering poetic characteristics
  --- can easily be incorporated into existing machine translation
  systems such as Moses \cite{moses}. The next thing to do with this
  research would be to actually implement these methods, and compare
  the results of the new system with existing systems. One could also
  look at how other state of the art systems (for example neural
  machine translation systems) could be integrated with these
  hypotheses. }
\paragraph{}{In conclusion: translating poetry is hard. Both manually
  and computationally, there are many pitfalls which must be avoided
  to ensure a translation is faithful to the original and fluently
  written. I hope that with this report, the rarely-explored field of
  computational translation of poetry may find some new ideas. Perhaps
  in the future we may be able to update the quote which the
  introduction started with:\\}

\begin{minipage}{\linewidth}
\begin{displayquote}
  Poetry is that which is lost in translation, \\
  Unless we use a computational calculation.  \\  
\end{displayquote}
\end{minipage}

% \section*{Notes on this version of the paper}
% \paragraph{}{The theory section requires the following changes: a
%   gloss of how one would evaluate poetry (computationally and
%   manually) and a glossary. }
% \paragraph{}{The discussion section requires a further discussion of
%   using text-to-speech methods to identify word stress, and a
%   discussion of future work and implementations of my hypotheses. }

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "Report"
%%% End:
