\paragraph{}{Translating poetry is hard. Without an intimate knowledge
  of the source and target language, a translation of any text will,
  without a doubt, be lacking. Translating poetry is even harder: and
  to do it computationally, almost impossible. This research project
  explores how we may improve current computational approaches to
  allow for better-quality translations of poetry.}
\paragraph{}{Why would we want to translate poetry? A couple of
  reasons come to mind, which specifically motivated my interest in
  this problem:}
\begin{itemize}
\item Foreign-language poems translated to English have historically
  built up the rich tapestry of English literature: where would we be
  without Dante, Virgil, Homer?
\item Machine translation could be used to translate poems to and from
  languages which have small, if any, translation communities.
\item A faithful translation requires the translator to have intimate
  familiarity with both the source and target languages: a computer
  could do one of these jobs, allowing for a lighter cognitive load on
  the translator and perhaps a better-quality translation.
\item Translated poems are often worth reading for their own merits,
  and it would be an interesting creative exercise to analyze machine
  translated poems.
\item The hope that a more computational, quantitative treatment of
  poetry and poetic analysis will encourage the more
  scientifically-minded to explore poetry and the beauty contained
  therein.
\end{itemize}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "Poster"
%%% End:
